Non-kissing and non-crossing complexes \\ for locally gentle algebras

### 2018-07-10 draft/accordionComplexSurfaces.tex
example Cambrian

\begin{example}[Cambrian paths]
\label{exm:CambrianPath}
\vincent{TODO: Cambrian paths}
\end{example}

@@

The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}) look like~\raisebox{-.1cm}{\includegraphics[scale=.3]{reversedPathsPM}}.
Moreover, $\bar Q$ has no infinite straight walk, while~$\bar Q\koszul$ has~$|Q_0|-1$ infinite straight walks.
Therefore, the corresponding surfaces have $1$ boundary component and genus~$0$.
The first few corresponding surfaces are represented in \fref{fig:reversedPathsSurfaces}.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{reversedPathsSurfaces3}}
	\caption{The surface~$\surface_{\bar Q}$ for the reversed path quivers with~$1, 2, 3$ vertices.}
	\label{fig:reversedPathsSurfaces}
\end{figure}

\end{example}

@@

The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}) look like~\raisebox{-.1cm}{\includegraphics[scale=.3]{doublePathsPModd}} when~$|Q_0|$ is odd and~\raisebox{-.1cm}{\includegraphics[scale=.3]{doublePathsPMeven}} when~$|Q_0|$ is even.
Moreover, $\bar Q$ and~$\bar Q\koszul$ have no infinite straight walks.
Therefore, the corresponding surfaces have no punctures and
\begin{itemize}
\item $1$ boundary component and genus~$(|Q_0|-1)/2$ when~$|Q_0|$ is odd, and 
\item $2$ boundary components and genus~$|Q_0|/2$ when~$|Q_0|$ is even.
\end{itemize}
The first few corresponding surfaces are represented in \fref{fig:doublePathsSurfaces}.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{doublePathsSurfaces}}
	\caption{The surface~$\surface_{\bar Q}$ for the double path quivers with~$1, 2, 3, 4$ vertices.}
	\label{fig:doublePathsSurfaces}
\end{figure}

\end{example}

@@

The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}) are empty since there is no blossom vertices.
Moreover, $\bar Q$ has~$1$ infinite straight walk when~$|Q_0|$ is odd and~$2$ when~$|Q_0|$ is even, and~$\bar Q\koszul$ always has~$2$ infinite straight walks.
Therefore, the corresponding surfaces have no boundary component and
\begin{itemize}
\item $3$ punctures and genus~$(|Q_0|-1)/2$ when~$|Q_0|$ is odd, and 
\item $4$ punctures and genus~$(|Q_0|-2)/2$ when~$|Q_0|$ is even.
\end{itemize}
The first few corresponding surfaces are represented in \fref{fig:doubleCyclesSurfaces}.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{doubleCyclesSurfaces}}
	\caption{The surface~$\surface_{\bar Q}$ for the double cycle quivers with~$3, 4, 5$ vertices.}
	\label{fig:doubleCyclesSurfaces}
\end{figure}

\end{example}

@@

\begin{proof}
Let~$c=\alpha_1\cdots\alpha_r$ be a non-trivial oriented cycle in some walk~$\omega$, with~$c, c^2\notin I$.
Assume that~$\omega$ is not of form~$\sigma c^{\pm\infty}$.
Then, up to cyclic reordering of the arrows in~$c$, it contains a substring of the form~$\beta^\pm c^l \alpha_1\cdots\alpha_j\gamma^\pm$, for some arrows~$\beta,\gamma$ not in~$c$, some~$l\geq 1$, and some~$j<r$.
Since~$\bar Q$ is locally gentle, this substring is~$\beta^{-1} c^l \alpha_1\cdots\alpha_j\gamma^{-1}$.
Hence, the walk~$\omega$ contains the two substrings~$\beta^{-1}\alpha_1\cdots\alpha_j\alpha_{j+1}$ and~$\alpha_r\alpha_1\cdots\alpha_j\gamma^{-1}$, and thus self-kisses.

@@

Once this cycle is completed, since~$\gamma$ cannot cross itself, it has to continue cycling infinitely around the puncture.  This proves the claim.
\vincent{Should we also give here a geometric argument?}
\end{proof}

@@

Thus~$F$ contains a curve going to the puncture.
\vincent{Should we also give here a geometric argument? Pierre-Guy, could you add that?}
\pierreguy{Done; I will need a picture, please, Vincent.}

\end{proof}

@@

\begin{definition}
The \defn{$\b{g}$-vector fan} of~$\bar Q$ is the collection of cones
\[
\gvectorFan \eqdef \bigset{\R_{\ge0} \gvectors{F}}{F \text{ non-kissing face of } \RNKC}.
\]
\end{definition}

>>>>

\begin{example}[Cambrian paths]
\label{exm:CambrianPath}
Consider a \defn{Cambrian quiver}, that is any orientation of a line with no relations (\aka type~$A$ quiver).
Two such quivers are represented in \fref{fig:CambrianPathsQuivers}.
Note that we can choose to position the blossom vertices in such a way that for any arrow~$\beta$, the arrows~$\alpha$ and~$\gamma$ such that~$\alpha\beta \in I$ and~$\beta\gamma \in I$ are on the right of~$\beta$.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.45]{CambrianPathsQuivers}}
	\caption{Two Cambrian quivers on~$5$ vertices.}
	\label{fig:CambrianPathsQuivers}
\end{figure}

The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}) form a cycle with green and red arrows alternating.
Moreover, $\bar Q$ and~$\bar Q\koszul$ have no infinite straight walks.
Since~$|Q_1| = |Q_0|-1$, the corresponding surfaces have $1$ boundary component and genus~$0$.
Two examples are represented in \fref{fig:reversedPathsSurfaces}.
Note that the dissection~$\dissection_{\bar Q}$ is a triangulation with no internal triangles, and the orientation of~$\bar Q$ indicates how to glue these triangles.
For instance, the Cambrian quiver completely oriented in one direction yields a fan triangulation, were all internal arcs are incident to the same vertex, as in \fref{fig:reversedPathsSurfaces}\,(left).

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{CambrianPathsSurfaces}}
	\caption{The surface~$\surface_{\bar Q}$ for two Cambrian paths on~$5$ vertices.}
	\label{fig:CambrianPathsSurfaces}
\end{figure}
\end{example}

@@

The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}) look like~\raisebox{-.1cm}{\includegraphics[scale=.3]{reversedPathsPM}}.
Moreover, $\bar Q$ has no infinite straight walk, while~$\bar Q\koszul$ has~$|Q_0|-1$ infinite straight walks.
Since~$|Q_1| = 2|Q_0|-2$, the corresponding surfaces have $1$ boundary component and genus~$0$.
The first few corresponding surfaces are represented in \fref{fig:reversedPathsSurfaces}.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{reversedPathsSurfaces3}}
	\caption{The surface~$\surface_{\bar Q}$ for the reversed path quivers with~$1, 2, 3$ vertices.}
	\label{fig:reversedPathsSurfaces}
\end{figure}
\end{example}

@@

The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}) look like~\raisebox{-.1cm}{\includegraphics[scale=.3]{doublePathsPModd}} when~$|Q_0|$ is odd and~\raisebox{-.1cm}{\includegraphics[scale=.3]{doublePathsPMeven}} when~$|Q_0|$ is even.
Moreover, $\bar Q$ and~$\bar Q\koszul$ have no infinite straight walks.
Since~$|Q_1| = 2|Q_0|-2$, the corresponding surfaces have no punctures and
\begin{itemize}
\item $1$ boundary component and genus~$(|Q_0|-1)/2$ when~$|Q_0|$ is odd, and 
\item $2$ boundary components and genus~$(|Q_0|-2)/2$ when~$|Q_0|$ is even.
\end{itemize}
The first few corresponding surfaces are represented in \fref{fig:doublePathsSurfaces}.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{doublePathsSurfaces}}
	\caption{The surface~$\surface_{\bar Q}$ for the double path quivers with~$1, 2, 3, 4$ vertices.}
	\label{fig:doublePathsSurfaces}
\end{figure}
\end{example}

@@

The corresponding perfect matchings (see Remark~\ref{rem:propertiesSurface}) are empty since there is no blossom vertices.
Moreover, $\bar Q$ has~$1$ infinite straight walk when~$|Q_0|$ is odd and~$2$ when~$|Q_0|$ is even, and~$\bar Q\koszul$ always has~$2$ infinite straight walks.
Since~$|Q_1| = 2|Q_0|$, the corresponding surfaces have no boundary component and
\begin{itemize}
\item $3$ punctures and genus~$(|Q_0|-1)/2$ when~$|Q_0|$ is odd, and 
\item $4$ punctures and genus~$(|Q_0|-2)/2$ when~$|Q_0|$ is even.
\end{itemize}
The first few corresponding surfaces are represented in \fref{fig:doubleCyclesSurfaces}.

\begin{figure}[h]
	\capstart
	\centerline{\includegraphics[scale=.7]{doubleCyclesSurfaces}}
	\caption{The surface~$\surface_{\bar Q}$ for the double cycle quivers with~$3, 4, 5$ vertices.}
	\label{fig:doubleCyclesSurfaces}
\end{figure}
\end{example}

@@

\para{Combinatorial proof}
Let~$c=\alpha_1\cdots\alpha_r$ be a non-trivial oriented cycle in some walk~$\omega$, with~$c, c^2\notin I$.
Assume that~$\omega$ is not of form~$\sigma c^{\pm\infty}$.
Then, up to cyclic reordering of the arrows in~$c$, it contains a substring of the form~$\beta^\pm c^l \alpha_1\cdots\alpha_j\gamma^\pm$, for some arrows~$\beta,\gamma$ not in~$c$, some~$l\geq 1$, and some~$j<r$.
Since~$\bar Q$ is locally gentle, this substring is~$\beta^{-1} c^l \alpha_1\cdots\alpha_j\gamma^{-1}$.
Hence, the walk~$\omega$ contains the two substrings~$\beta^{-1}\alpha_1\cdots\alpha_j\alpha_{j+1}$ and~$\alpha_r\alpha_1\cdots\alpha_j\gamma^{-1}$, and thus self-kisses.

@@

Once this cycle is completed, since~$\gamma$ cannot cross itself, it has to continue cycling infinitely around the puncture.  This proves the claim.
\end{proof}

@@

Thus~$F$ contains a curve going to the puncture.
\end{proof}

@@

\begin{definition}
\vincent{Mettre la def dans les thm}
The \defn{$\b{g}$-vector fan} of~$\bar Q$ is the collection of cones
\[
\gvectorFan \eqdef \bigset{\R_{\ge0} \gvectors{F}}{F \text{ non-kissing face of } \RNKC}.
\]
\end{definition}
